Extending Morphological Chains with Supervised Learning and Cross-Language Features

### 2015-12-10 paper/paper.tex
learning stuff

\begin{document}

\title{Extending Morphological Chains with Supervised Learning and Cross-Language Features}

@@

\section{Background Work}
Our model builds on work done by Narasimhan et al. (2015) in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner. The idea behind morphological chains is that complex words are constructed by attaching morphemes simpler words. Thus, complex words have words from which they are directly derived, which we call \emph{parent} words. Likewise, the derived word is a \emph{child} word. A morphological chain is a sequence of words such that consecutive pairs form a parent-child relationship. For example, the word \emph{unsustainable} can be constructed from \emph{sustain}, as demonstrated in the chain \emph{sustain} $\rightarrow$ \emph{sustainable} $\rightarrow$ \emph{unsustainable}. Words without parents are called \emph{base} words. In our example, \emph{sustain} is a base word.

@@

Note that parent words usually undergo changes only when attaching suffixes. Thus there is one \emph{Prefix} class, but there are a variety of suffix classes. When acquiring a suffix, the parent word can:
\begin{enumerate}
    \item undergo no change (\emph{class} $\rightarrow$ \emph{classes})
    \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
    \item delete a character (\emph{believe} $\rightarrow$ \emph{believing})
    \item modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
\end{enumerate}
These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}). Finally, there is a \emph{Stop} type, which is used to signify that the word is a base word.

@@

In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents. In generating our vectors, we use text from Wikipedia,
to generate the english word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.

@@

@@

\section{Using Cross-Language Features and Candidates}
Due to the simple nature of our parent candidate generation routine,
there are many cases in which the potential parent is not actually generated;
for example, the morphological parent of "feet" is the singular "foot";
however, there is no prefix or suffix that turns one into the other, and
therefore the parent for "feet" is not generated by our system.

However, other languages may not have such an irregular transformation
for the same words, for example in Turkish the corresponding translations
for "feet" and "foot" are "ayaklar" and "ayak". A properly trained Turkish
morphological parent model will correctly detect "ayak" as a parent of "ayaklar",
and can be used to suggest morphological parents in English after translation.

>>>>

\begin{document} 
\title{Extending Morphological Chains with Supervised Learning and Cross-Language Features}

@@

\section{Unsupervised Model}
Our model builds on work done by Narasimhan et al. (2015) in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner. The idea behind morphological chains is that complex words are constructed by attaching morphemes simpler words. Thus, complex words have words from which they are directly derived, which we call \emph{parent} words. Likewise, the derived word is a \emph{child} word. A morphological chain is a sequence of words such that consecutive pairs form a parent-child relationship. For example, the word \emph{unsustainable} can be constructed from \emph{sustain}, as demonstrated in the chain \emph{sustain} $\rightarrow$ \emph{sustainable} $\rightarrow$ \emph{unsustainable}. Words without parents are called \emph{base} words. In our example, \emph{sustain} is a base word.

@@

Note that parent words usually undergo changes only when attaching suffixes. Thus there is one \emph{Prefix} class, but there are a variety of suffix classes. When acquiring a suffix, the parent word can: (1) undergo no change (\emph{class} $\rightarrow$ \emph{classes}), (2) repeat a character (\emph{star} $\rightarrow$ \emph{starring}), (3) delete a character (\emph{believe} $\rightarrow$ \emph{believing}), (4) modify a character (\emph{parry} $\rightarrow$ \emph{parried}). These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}). Finally, there is a \emph{Stop} type, which is used to signify that the word is a base word.

@@

In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents. In generating our vectors, we use text from Wikipedia along with a corpus of news articles.

@@

\subsection{Unsupervised Learning}
Recall that in our model, we want to find weights $\theta$ for the feature vectors $\phi(w,z)$, where $w\in\mathcal W$ and $z\in\mathcal Z$. Define the probabilities of a word-candidate pair $(w,z)$ as $P(w,z)\propto e^{\theta\cdot\phi(w,z)}$. Then the probability of a candidate $z$ ocuring given $w$ is
\begin{equation}
    P(z|w) = \frac{e^{\theta\cdot\phi(w,z)}} {\sum_{z'\in C(w)} e^{\theta\cdot\phi(w,z')}}
\end{equation}

Let our set of observed words be $D$. We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$. We use the EM-algorithm to maximize the likelihood over $\theta$ as given by

\begin{equation}
    \begin{split}
        L(\theta; D) &= \prod_{w^*\in D} P(w^*) \\
        &= \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
        &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
    \end{split}
\end{equation}

We cannot compute $L(\theta; D)$ directly since we cannot compute over all strings in $\Sigma^*$. Instead, we approximate this distribution by considering only strings which are similar to those encountered in $D$.

We use the method of Contrastive Estimation (Smith and Eisner, 2005) and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$. Toward this end, we transpose pairs of consecutive letters of $w$ near both ends of the word. We also do both simultaneously. Together these form our set of neighbors for $w$. The neighbors form a proxy for $\Sigma^*$, and represent the set of strings we want to reduce the probability of seeing in our model because they are not observed. This has the benefit of providing contrast in the structure of words while not requiring the model to look at the entire $\Sigma^*$.

With this substitution, we can formulate the contrastive likelihood as
\begin{equation}
    L_C(\theta; D) = \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
\end{equation}

@@

\section{Using Cross-Language Features and Candidates}
Due to the simple nature of our parent candidate generation routine,
there are many cases in which the potential parent is not actually generated;
for example, the morphological parent of ``feet'' is the singular ``foot'';
however, there is no prefix or suffix that turns one into the other, and
therefore the parent for ``feet'' is not generated by our system.

However, other languages may not have such an irregular transformation
for the same words, for example in Turkish the corresponding translations
for ``feet'' and ``foot'' are ``ayaklar'' and ``ayak''. A properly trained Turkish
morphological parent model will correctly detect ``ayak'' as a parent of ``ayaklar'',
and can be used to suggest morphological parents in English after translation.
